Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks

### 2023-02-06 paper-1/manuscript.tex
Fixed TA refs

Neuronal plasticity, on the other hand, has largely been neglected in previous work, used as a mere afterthought in the neuron model \citep{Mill2011-ah} or rejected a priori as incapable of producing DD \citep{Yarden2017-eh}. However, neuronal adaptation, i.e., an experience-dependent reduction in intrinsic neuronal sensitivity independent of synaptic function, is a common phenomenon in excitatory neurons \citep{Sanchez-Vives2000-df, Henze2001-xd, Sanchez-Aguilera2014-fd, Teeter2018-iz} that has been successfully used to model stimulus-specific adaptation (SSA), a precursor to and prerequisite for DD \citep{Garagnani2011-eu}. Neuronal adaptation is well placed to play a role in DD: It establishes quickly and recovers over the course of up to several seconds \citep{Henze2001-xd, Mensi2012-au}, allowing it to act as a memory of recent activity in a neuron.

@@

When a neuron's membrane potential reached the firing threshold $V_\theta$, a spike was emitted, and the potential was clamped to $V_m = V_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons). Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $V_\theta = \theta_0 = -54$~mV \citep{Mensi2012-au}. In contrast, excitatory neurons were modeled with threshold adaptation (TA) as in \cite{Teeter2018-iz}, following
\begin{align}
    V_\theta &= \theta_0 + \theta(t) \nonumber \\
    \tau_{\theta} \frac{d\theta}{dt} &= -\theta + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
\end{align}
with increment $\hat{\theta} = 1$~mV and decay time constant $\tau_{\theta} = 1$~s \citep{Pozzorini2015-ei}.

>>>>

Neuronal plasticity, on the other hand, has largely been neglected in previous work, used as a mere afterthought in the neuron model \citep{Mill2011-ah} or rejected a priori as incapable of producing DD \citep{Yarden2017-eh}. However, neuronal adaptation, i.e., an experience-dependent reduction in intrinsic neuronal sensitivity independent of synaptic function, is a common phenomenon in excitatory neurons \citep{Sanchez-Vives2000-df, Henze2001-xd, Sanchez-Aguilera2014-fd, Teeter2018-iz} that has been successfully used to model stimulus-specific adaptation (SSA), a precursor to and prerequisite for DD \citep{Garagnani2011-eu}. Neuronal adaptation is well placed to play a role in DD: It establishes quickly and recovers over the course of up to several seconds \citep{Henze2001-xd, Pozzorini2015-ei}, allowing it to act as a memory of recent activity in a neuron.

@@

When a neuron's membrane potential reached the firing threshold $V_\theta$, a spike was emitted, and the potential was clamped to $V_m = V_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons). Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $V_\theta = \theta_0 = -54$~mV \citep{Mensi2012-au}. In contrast, excitatory neurons were modeled with threshold adaptation (TA) as in \cite{Teeter2018-iz}, following
\begin{align}
    V_\theta &= \theta_0 + \theta(t) \nonumber \\
    \tau_{\theta} \frac{d\theta}{dt} &= -\theta + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
\end{align}
with increment $\hat{\theta} = 1$~mV and decay time constant $\tau_{\theta} = 1$~s \citep{Henze2001-xd, Pozzorini2015-ei}.
